---
title: "SF-crime"
output: html_document
---

[San Francisco crime dataset from Kaggle](https://www.kaggle.com/c/sf-crime)

## Data fields

* Dates - timestamp of the crime incident
* Category - category of the crime incident (only in train.csv). This is the target variable you are going to predict.
* Descript - detailed description of the crime incident (only in train.csv)
* DayOfWeek - the day of the week
* PdDistrict - name of the Police Department District
* Resolution - how the crime incident was resolved (only in train.csv)
* Address - the approximate street address of the crime incident
* X - Longitude
* Y - Latitude


```{r, echo=FALSE, message=FALSE}
library(ggplot2)    # Data visualization
library(data.table) # Much faster data reading
library(dplyr)      # Data aggregation etc.
library(scales)     # Plot scaling
library(ggmap)
```

```{r, echo = FALSE}
start.time <- Sys.time()
train <- fread("xzcat ../data/train.csv.xz", header = TRUE)
end.time <- Sys.time()

cat(sprintf("Loaded %d rows in %.1f seconds.\n", nrow(train), end.time - start.time))
```

```{r, echo = FALSE}
train$Dates <- as.POSIXct(train$Dates, "PST")
train$Year  <- as.numeric(format(train$Dates, "%Y"))
train$Month <- as.numeric(format(train$Dates, "%m"))

train$Category <- as.factor(train$Category)
```

### Frequency of Events per Category

```{r, echo = FALSE}
# compute montly count per Category
train.agg <- train %>% group_by(Year, Month, Category) %>% summarize(count=n())
train.agg$Date <- as.POSIXct(paste(train.agg$Year, train.agg$Month, "01", sep = "-")) # set date to first of month

# set monthly count to 0 for missing values
alldates <- data.frame(Date=with(train.agg, seq(min(Date), max(Date), by="month")))
allcatdates <- merge(alldates, data.frame(Category=levels(train$Category)))
train.agg <- merge(train.agg, allcatdates, by=c("Date", "Category"), all = TRUE)
train.agg[is.na(train.agg$count)]$count <- 0

# plot montly count per Category
breaks <- seq(as.POSIXct("2003-01-01"), as.POSIXct("2015-01-01"), by="2 years")
ggplot(train.agg, aes(x = Date, y = count, group = Category, col = Category)) +
  geom_line(size=0.2) +
  facet_wrap(~Category, ncol = 5, scales = "free_y") +
  theme(legend.position = "none",
        axis.text = element_text(size = 5),
        strip.text = element_text(size = 5)) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_datetime(breaks = breaks, labels = format(breaks, "%Y")) +
  labs(title="Monthly Frequency of Crime Events per Category")
```

### Location of Events in 2010

```{r, echo = FALSE, message=FALSE}
train <- subset(train, Y < 40)

train <- subset(train, Year == 2010)

#gglMap <- get_map("San Francisco", zoom = 12, color = "bw")
osmMap<-get_map("San Francisco", zoom = 12, source="osm", color = "bw")

ggmap(osmMap) +
  geom_point(data = train, aes(x = X, y = Y, col = PdDistrict), size = 0.1, alpha = 0.3) +
  facet_wrap(~PdDistrict, ncol = 5) +
  theme(legend.position = "none", axis.text = element_text(size = 5))
```
